---
title: "R Notebook"
output: html_notebook
---

#Linear regression Model to find the amount of influence on profit by other parameters for a startup company.

load the file 50-startups.csv and save it as startup dataframe.

```{r}
View(startup)
head(startup)
dim(startup)
rownames(startup)
colnames(startup)
str(startup)
summary(startup)
```

#Let's start with Exploratory data analysis now

```{r}
attach(startup)
hist(R.D.Spend)
```

# we can observe that between 60000 to 80000 more companies are investing that much of amount for research and development

```{r}
Profit
cor(Profit,R.D.Spend) # we get 0.9729005 which is positive correlation between the profit and R.D.Spend
```


# Visualisation of variables
```{r}
plot(R.D.Spend,Profit) # here R.D.Spend is on x axis independent variable and Profit on y axis as it is dependent variable
# here profit depends on R.D.Spend
plot(Administration,Profit)
cor(Administration,Profit)
plot(Profit,Administration)

# always plot profit as dependent variable on y-axis and independent variables on x-axis

plot(Marketing.Spend,Profit)

plot(State,Profit)

print(State)

str(startup) # only state is in factor

```

# In linear regression model, we can only get insights for numerical startup type quantities but not for textual so we need to convert the factor to numeric
# It means we need to convert narrow format of state column to wide format. we can do that simply by considering dummies.


```{r}
startup$DummyNY <- ifelse(State=="New York",1,0)
startup$DummyNY

head(startup)

str(startup)

# we can remove State column as we got dummyNY column

startup$State <- NULL

head(startup)
pairs(startup)

library(corrgram)
corrgram(startup)

library(GGally)
ggpairs(startup)

```

# Model Building

```{r}
multReg <- lm(Profit~R.D.Spend+Administration+Marketing.Spend+DummyNY,startup)
summary(multReg)
# Multiple R-squared:  0.951,	Adjusted R-squared:  0.9467
```
# Remove variables with highest p-value(>0.05)
# Remove Administration variable and check for R-squared value and adjusted R-squared value 
```{r}
multReg1 <- lm(Profit~R.D.Spend+Marketing.Spend+DummyNY,startup)
summary(multReg1)
# Multiple R-squared:  0.9508,	Adjusted R-squared:  0.9476
```
# Remove variable with high p-value(>0.05)
# Remove DummyNY
```{r}
multReg2 <- lm(Profit~R.D.Spend+Marketing.Spend,startup)
summary(multReg2)
# Multiple R-squared:  0.9505,	Adjusted R-squared:  0.9483
```
# Now let's remove Marketing.Spend and check R-squared and adjusted R-squared values
```{r}
multReg3 <- lm(Profit~R.D.Spend,startup)
summary(multReg3)
# Multiple R-squared:  0.9465,	Adjusted R-squared:  0.9454 
# observe the values of R-squared and adjusted R-squared
# R-squared decreases with decrease in number of independent variables, so more independent variables more R-Squared
# Adjusted R-squared value initially increases but finally decreases
# fit the model with high adjusted R-Squared 
```
# we can see that R-squared and Adjusted R-squared values dropped significantly so we can't neglect Marketing.Spend variable as it has p-value only slightly higher.

# So Final model:-
```{r}
multReg <- lm(Profit~R.D.Spend+Marketing.Spend,startup)
summary(multReg)
```
# Visualisation of final model
```{r}
plot(multReg)
```

