---
title: "R Notebook"
output: html_notebook
---

#Here we have a data from a bank and they want to know the customers that are likely default on payment.

#Load the file 'Log-Reg-Case-Study.csv' and save it as data

```{r}
View(data)
head(data)
```

```{r}
summary(data)
str(data)
```
```{r}
sum(is.na(data))
```

# Exploratory Data Analysis
# Univariate Analysis

```{r}
summary(data$Age)
boxplot(data$Age)
quantile(data$Age,0.995) # 99.5th percentile value
quantile(data$Age,0.999) # 99.9th percentile value
max(data$Age)
```

# so in 99.9th percentile the age is 75 and in 0.001 somewhere it has maximum age as 85
# Let's do outlier capping
# if condition age >75 cap it to 75
```{r}
data$Age <- ifelse(data$Age>75,75,data$Age)
summary(data$Age)
max(data$Age)
```
# adjusted maximum value in age so as to prevent the outlier problem

# now lets fill the blanks or missing values in categorical variables like housing
# if the missing value is in categorical variable we need to replace it with mode
# if the missing value is in numerical variable replace it with mean/median appropriately

```{r}
summary(data$Housing)
# 9 blanks here 
# check for mode we can see that A152 has higher mode so lets replace blanks with it
```
```{r}
data$Housing[data$Housing==""]<-"A152"
summary(data$Housing)
# all blanks are replaced 
```

# Bivariate Analysis

```{r}
summary(data$Num_Dependents)
table(data$Num_Dependents,data$Default_On_Payment)
aggregate(data$Num_Dependents,by=list(data$Num_Dependents,data$Default_On_Payment),length)


# missing value imputation ( BiVariant analysis)
table(data$Job_Status,data$Default_On_Payment)
levels(data$Job_Status)
aggregate(data$Job_Status,by=list(data$Job_Status,data$Default_On_Payment),length)
summary(data$Job_Status)
```
```{r}
# now replace the blanks with highest percentage of category rather than with absolute values 
# we used absolute values i mean mode for univarient analysis

# for Bivarient analysis we need to go with highest percentage rather than by mode
# here for job status the highest percentage for default is A174
#so lets replace blanks with that value of A174

data$Job_Status[data$Job_Status==""]<-"A174"
summary(data$Job_Status)
str(data$Job_Status)
```
# now all the categorical independent variable have to be converted into numeric so that we can plot or to create model

# Grouping for dummy variable creation
# here dummmy variable converts categorical data to numeric
# if there are 10 independent category variables 9 dummy variables are needed generally but we can group the varaiables to get few dummy variables for easy analysis

# Dummy variables for job_status
```{r}
summary(data$Job_Status)

# we can see there are 4 variablesand 3 dummy Variables as below:-
data$Dummy_Job_Status_A171 <- ifelse(data$Job_Status=="A171",1,0)
data$Dummy_Job_Status_A172 <- ifelse(data$Job_Status=="A172",1,0)
data$Dummy_Job_Status_A173 <- ifelse(data$Job_Status=="A173",1,0)
```
# Dummy Variables for Purpose_Credit_taken
```{r}
table(data$Purpose_Credit_Taken,data$Default_On_Payment)
# 10 categorial levels here we actually need to create 9 dummy variables but instead on observation we can make it into groups as below:-
# but we are here classifying it as high,low

data$Dummy_Purpose_Credit_Taken_Low <- ifelse(data$Purpose_Credit_Taken=="P41"| data$Purpose_Credit_Taken=="P43"|data$Purpose_Credit_Taken=="P48",1,0)
data$Dummy_Purpose_Credit_Taken_High <- ifelse(data$Purpose_Credit_Taken=="P49"|data$Purpose_Credit_Taken=="P40"|
                                                      data$Purpose_Credit_Taken=="P45"|data$Purpose_Credit_Taken=="P50"|data$Purpose_Credit_Taken=="P46",1,0)
```
# Dummy variable for status checking account :- 2 variables high,medium
```{r}
table(data$Status_Checking_Accnt,data$Default_On_Payment)
data$Dummy_Status_Checking_Accnt_High <- ifelse(data$Status_Checking_Accnt=="S11",1,0)
data$Dummy_Status_Checking_Accnt_Medium <- ifelse(data$Status_Checking_Accnt=="S12",1,0)
summary(data$Status_Checking_Accnt)
```

# Dummy Variables for Credit_History:- high or low
```{r}
table(data$Credit_History,data$Default_On_Payment)
# 5 categorical variables
data$Dummy_Credit_History_High <- ifelse(data$Credit_History=="A30"|data$Credit_History=="A31",1,0)
data$Dummy_Credit_History_Low <- ifelse(data$Credit_History=="A34",1,0)
```
# Dummy Variables for Years_At_Present_Employment :- high or medium
```{r}
table(data$Years_At_Present_Employment,data$Default_On_Payment)
data$Dummy_Years_At_Present_Employment_High <- ifelse(data$Years_At_Present_Employment=="E71"| data$Years_At_Present_Employment=="E72",1,0)
data$Dummy_Years_At_Present_Employment_Medium <- ifelse(data$Years_At_Present_Employment=="E73",1,0)
```
# Dummy variables for Marital_Status_Gender
```{r}
table(data$Marital_Status_Gender,data$Default_On_Payment)
data$Dummy_Marital_status_Gender <- ifelse(data$Marital_Status_Gender=="A91"|data$Marital_Status_Gender=="A92",1,0)
```
# Dummy variable for Other_Debtors_Guarantors
```{r}
table(data$Other_Debtors_Guarantors,data$Default_On_Payment)
data$Dummy_Other_Debtors_Guarantors <- ifelse(data$Other_Debtors_Guarantors=="A103",0,1)
# note here we took 0,1 instead of 1,0
```

# Dummy Variable for housing
```{r}
summary(data$Housing)
data$Dummy_Housing <- ifelse(data$Housing=="A152",0,1)
```
# Dummy Variable for Foreign worker
```{r}
data$Dummy_Foreign_Worker <- ifelse(data$Foreign_Worker=="A201",1,0)
```

# convert numeric varibles to dummy variables (Grouping numeric variables like age)
# if age is less than 30 take it as default on payment
```{r}
table(data$Age,data$Default_On_Payment)
data$Dummy_Age_Group <- ifelse(data$Age<30,1,0)
```
# Dummy Variable for credit_amount
# if the credit amount is less than 5000 take it as not defaulted in payment
```{r}
data$Dummy_Credit_Amount <- ifelse(data$Credit_Amount<5000,0,1)
```
# Dummy variable for current address in years
```{r}
data$Dummy_Current_Address_Yrs <- ifelse(data$Current_Address_Yrs==1,0,1)
```
# Num_credits and Num_Dependents do not show significant relationship with the dependent variable
```{r}
View(data)
```

# Model Building
# Split data into 70% train and 30% test

```{r}
set.seed(123) # to reproduce the same sample again
train=sample(1:nrow(data),nrow(data)*0.7)
test=-train
ModelData=data[train,]
head(ModelData)

# 70% of total data indices we have selected using train is placed in ModelData now
nrow(ModelData)

# to assign remaining 30% data to VaqlidationData can be done as below
# method 1:-
ValidationData=data[test,]
nrow(ValidationData)
```
# Model Building
```{r}
library('MASS')
library(car)
#lm stands for linear model
# all variables to be analysed are to be included in the equation,with dataset's name at the end
model_1 <- glm(Default_On_Payment~Dummy_Job_Status_A171+Dummy_Job_Status_A172+Dummy_Job_Status_A173+Dummy_Purpose_Credit_Taken_High+Dummy_Purpose_Credit_Taken_Low+
                   Dummy_Status_Checking_Accnt_High+Dummy_Status_Checking_Accnt_Medium+Dummy_Credit_History_High+Dummy_Credit_History_Low+Dummy_Years_At_Present_Employment_High+
                   Dummy_Years_At_Present_Employment_Medium+Dummy_Marital_status_Gender+Dummy_Other_Debtors_Guarantors+Dummy_Housing+Dummy_Foreign_Worker+Dummy_Age_Group+
                   Dummy_Credit_Amount+Dummy_Current_Address_Yrs+Duration_in_Months,family=binomial(logit),data=ModelData)
summary(model_1)
vif(model_1)
```

# vif value should be less than 2. here we can see that there are 2 variables with vif > 2 so we need to remove one variable
# here vif() output is high for 2 variables Dummy_Job_Status_A172 and Dummy_Job_Status_A173
# so one variable will be removed

# perform regression on both variable and remove the model with less R-squared value
```{r}
Multicoll_out1 <- lm(Default_On_Payment~Dummy_Job_Status_A172,data=ModelData)
summary(Multicoll_out1)

Multicoll_out2 <- lm(Default_On_Payment~Dummy_Job_Status_A173,data=ModelData)
summary(Multicoll_out2)
# so now remove Dummy_Job_Status_A173 and repeat vif
```

```{r}
model_2<- glm(Default_On_Payment~Dummy_Job_Status_A171+Dummy_Job_Status_A172+Dummy_Purpose_Credit_Taken_High+Dummy_Purpose_Credit_Taken_Low+
                   Dummy_Status_Checking_Accnt_High+Dummy_Status_Checking_Accnt_Medium+Dummy_Credit_History_High+Dummy_Credit_History_Low+Dummy_Years_At_Present_Employment_High+
                   Dummy_Years_At_Present_Employment_Medium+Dummy_Marital_status_Gender+Dummy_Other_Debtors_Guarantors+Dummy_Housing+Dummy_Foreign_Worker+Dummy_Age_Group+
                   Dummy_Credit_Amount+Dummy_Current_Address_Yrs+Duration_in_Months,family=binomial(logit),data=ModelData)
summary(model_2)
vif(model_2)
```

# All the variables have vif < 2. Now eleminate insignificant variables (p-value > 0.05) one by one

# Remove Dummy_Job_Status_A172 as it has highest P-value
```{r}
model_3 <-glm(Default_On_Payment ~ Dummy_Job_Status_A171+ 
    Dummy_Purpose_Credit_Taken_High + Dummy_Purpose_Credit_Taken_Low + 
    Dummy_Status_Checking_Accnt_High + Dummy_Status_Checking_Accnt_Medium + 
    Dummy_Credit_History_High + Dummy_Credit_History_Low + Dummy_Years_At_Present_Employment_High + 
    Dummy_Years_At_Present_Employment_Medium + Dummy_Marital_status_Gender + 
    Dummy_Other_Debtors_Guarantors + Dummy_Housing + Dummy_Foreign_Worker + 
    Dummy_Age_Group + Dummy_Credit_Amount + Dummy_Current_Address_Yrs + 
    Duration_in_Months, family = binomial(logit), data = ModelData)
summary(model_3)
```

# Remove Dummy_Credit_Amount as it is highly insignificant
```{r}
model_4 <-glm(Default_On_Payment ~ Dummy_Job_Status_A171+ 
    Dummy_Purpose_Credit_Taken_High + Dummy_Purpose_Credit_Taken_Low + 
    Dummy_Status_Checking_Accnt_High + Dummy_Status_Checking_Accnt_Medium + 
    Dummy_Credit_History_High + Dummy_Credit_History_Low + Dummy_Years_At_Present_Employment_High + 
    Dummy_Years_At_Present_Employment_Medium + Dummy_Marital_status_Gender + 
    Dummy_Other_Debtors_Guarantors + Dummy_Housing + Dummy_Foreign_Worker + 
    Dummy_Age_Group + Dummy_Current_Address_Yrs + 
    Duration_in_Months, family = binomial(logit), data = ModelData)
summary(model_4)
```

# Remove Dummy_Purpose_Credit_Taken_Low as it is insignificant
```{r}
model_5 <-glm(Default_On_Payment ~ Dummy_Job_Status_A171+ 
    Dummy_Purpose_Credit_Taken_High + 
    Dummy_Status_Checking_Accnt_High + Dummy_Status_Checking_Accnt_Medium + 
    Dummy_Credit_History_High + Dummy_Credit_History_Low + Dummy_Years_At_Present_Employment_High + 
    Dummy_Years_At_Present_Employment_Medium + Dummy_Marital_status_Gender + 
    Dummy_Other_Debtors_Guarantors + Dummy_Housing + Dummy_Foreign_Worker + 
    Dummy_Age_Group + Dummy_Current_Address_Yrs + 
    Duration_in_Months, family = binomial(logit), data = ModelData)
summary(model_5)
```

# Model Validation

```{r}
ValidationData$Predicted_Probability <- predict(model_5,ValidationData[,-17],type="response")
ValidationData$Predicted_Probability<-round(ValidationData$Predicted_Probability,2)
View(ValidationData)
```

# create the final flag-prediction based on probability for model and validation datasets
# Here if the probability is greater than 30% then consider that the customer defaults on payment
# threshold value is given by the bank.
```{r}
ValidationData$Final_Prediction <- ifelse(ValidationData$Predicted_Probability>0.30,1,0)
table(ValidationData$Default_On_Payment,ValidationData$Final_Prediction)
conf <- confusionMatrix(ValidationData$Final_Prediction,ValidationData$Default_On_Payment)
conf
```
# Final accuracy is 72%
# Sensitivity is 72.34%
# Specificity is 71.28% for cut-off=0.30
